Effects of epidemiological structure on the transient evolution of HIV virulence

### 2017-01-10 ms/ms.tex
edit figure captions

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig1.pdf}
\caption{{\bf Baseline dynamics.}
Time series of mean population \Lspvl. (a) Contrast between the three-stage Shirreff model and the single-stage model calibrated to varying initial exponential growth rates, $r$. (b) Effects of varying initial infectious density $I(0)$. (c) Effects of varying initial mean virulence $\alpha(0)$. The $r=0.042$ (orange, dotted) curve in panel (a), calibrated to match the epidemic dynamics of Shirreff \etal's model \cite{shirreff_transmission_2011}, shows that our simplified model can produce similar virulence trajectories. Panels b and c illustrate the sensitivity of virulence trajectories to initial conditions $I(0)$ and $\alpha(0)$, which we hold constant in our simulations.}
\label{fig:panel3}
\end{figure}

@@

\begin{table}[h!]
\caption{Parameter ranges/values.  Values of $c$ and $\rho$ are doubled from those given by Champredon \etal\ because we keep track of individuals in the model, while they keep track of couples. Starred (*) parameters (used in \figurename~\ref{fig:panel3}), and descriptions of Hill function coefficients, are taken from \cite{shirreff_transmission_2011}.}
\centering
\begin{tabular}{c p{2in} c l}
\hline 
Notation & Description & Range/Value & Source\\
\hline % inserts single horizontal line
$\rho$ & Partnership formation rate & 1/10-2/5 per year & \cite{champredon_hiv_2013} \\
$c$ & Partnership dissolution rate & 1/15-1/5 (1.25*) per year & \cite{champredon_hiv_2013} \\
$c_u/c_w$ & Relative contact rate for uncoupled transmission & 1/5-5 & Assumption \\
$c_e/c_w$ & Relative contact rate extra-couple & 0.01-1 & \cite{champredon_hiv_2013} \\
$\beta_P$ & Rate of transmission during primary infection & 1.31-5.09 (2.76*) per year & \cite{hollingsworth_hiv1_2008} \\
$\beta_D$ & Rate of transmission during high transmission disease stage & 0.413-1.28 (0.76*) per year & \cite{hollingsworth_hiv1_2008} \\
$D_P$ & Duration of primary infection & 1.23/12-6/12 (0.25*) years & \cite{hollingsworth_hiv1_2008} \\
$D_D$ & Duration of high transmission disease stage & 4.81/12-14/12 (0.75*) years & \cite{hollingsworth_hiv1_2008} \\
$\tsub{\beta}{max}$ & Maximum rate of transmission during asymptomatic stage & 0.317 per year & \cite{shirreff_transmission_2011} \\
$\beta_{50}$ & SPVL at which infectiousness is half maximum & 13938 copies per ml & \cite{shirreff_transmission_2011} \\
$\beta_k$ & Hill coefficient: steepness of increase in infectiousness as a function of SPVL & 1.02 & \cite{shirreff_transmission_2011} \\
$\tsub{D}{max}$ & Duration of primary infection & 25.4 years & \cite{shirreff_transmission_2011} \\
$D_{50}$ & SPVL at which duration of asymptomatic infection is half maximum & 3058 copies per ml & \cite{shirreff_transmission_2011} \\
$D_{k}$ & Hill coefficient: steepness of decrease in duration as a function of SPVL & 0.41 & \cite{shirreff_transmission_2011} \\
$\sigma_M$ & Mutation standard deviation of $\log_{10}$ SPVL & 0.12 & \cite{shirreff_transmission_2011} \\
$\tsub{\alpha}{min}$ & Minimum $\log_{10}$ SPVL & 2 & \cite{shirreff_transmission_2011}\\
$\tsub{\alpha}{max}$ & Maximum $\log_{10}$ SPVL & 7 & \cite{shirreff_transmission_2011}\\
$n$ & Number of strains & 21 (51*) & Assumption\\
$\mu$ & Mean number of non-cohabiting sexual partners & 0.103 - 1.206 & \cite{omori2015dynamics}\\
$\kappa$ & Squared coefficient of variation of number of non-cohabiting sexual partners & 0.01 - 100 & Assumption\\
\hline
\end{tabular}
\label{table:parmsTable}
\end{table}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig2.pdf}
\caption{{\bf Envelopes of virulence trajectories (expected
time of progression to AIDS) under all models.}
All models were run until $t=4000~\textrm{years}$; truncated series are shown here.}
\label{fig:virtraj}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig3.pdf}
\caption{{\bf Univariate distributions of summary statistics.}
The distribution of equilibrium expected progression
time (lower left panel) for the random mixing model is very narrow, and has been replaced by a point in order to preserve the vertical axis scaling.}
\label{fig:unidist}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig4.pdf}
\caption{{\bf Pairs plot: bivariate relationships among summary statistics for each model structure.}
Dashed line in equilibrium vs. peak virulence plot shows 1:1 line. 100 values were sampled from each model to allow for clearer distinction between the models}
\label{fig:pairplot}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig5.pdf}
\caption{{\bf Sensitivity plot.}
For most parameters in the Latin hypercube sample and each summary statistic, the figure shows the distribution (points) and trend (smooth line) of the summary statistic as a function of the \emph{unscaled} parameter value, i.e. prior to adjusting the parameters to achieve the standard initial epidemic growth rate.}
\label{fig:plot_sens}
\end{figure}

>>>>

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig1.pdf}
\caption{{\bf Baseline dynamics.}
Time series of mean population \Lspvl\ using baseline parameters. Each initial condition was varied independently in order to test their effects on transient dynamics of virulence evolution. Note that the initial condition has negligible effect on equilibrium virulence.
(a) Contrast between the three-stage Shirreff model and the single-stage model calibrated to varying initial exponential growth rates, $r$. achieved quicker.
(b) Effects of varying initial infectious density $I(0)$.
(c) Effects of varying initial mean virulence $\alpha(0)$.
The $r=0.042$ (orange, dotted) curve in panel (a), calibrated to match the epidemic dynamics of Shirreff \etal's model \cite{shirreff_transmission_2011}, shows that our simplified model can produce similar virulence trajectories. 
Although virulence dynamics are sensitive to initial conditions, we are still able to achieve similar qualitative dynamics dynamics across a range of parameters. 
For fair comparison, we hold them constant throughout our simulations.}
\label{fig:panel3}
\end{figure}

@@

\begin{table}[h!]
\caption{Parameter ranges/values. Parameters with fixed values are kept constant throughout simulations while other parameter values are taken from Latin hypercube samples using ranges specified in the table. Values of $c$ and $\rho$ are doubled from those given by Champredon \etal\ because we keep track of 	individuals in the model, while they keep track of couples. Starred (*) parameters (used in \figurename~\ref{fig:panel3}), and descriptions of Hill function coefficients, are taken from \cite{shirreff_transmission_2011}.}
\centering
\begin{tabular}{c p{2in} c l}
\hline 
Notation & Description & Range/Value & Source\\
\hline % inserts single horizontal line
$\rho$ & Partnership formation rate & 1/10-2/5 per year & \cite{champredon_hiv_2013} \\
$c$ & Partnership dissolution rate & 1/15-1/5 (1.25*) per year & \cite{champredon_hiv_2013} \\
$c_u/c_w$ & Relative contact rate for uncoupled transmission & 1/5-5 & Assumption \\
$c_e/c_w$ & Relative contact rate extra-couple & 0.01-1 & \cite{champredon_hiv_2013} \\
$\beta_P$ & Rate of transmission during primary infection & 1.31-5.09 (2.76*) per year & \cite{hollingsworth_hiv1_2008} \\
$\beta_D$ & Rate of transmission during high transmission disease stage & 0.413-1.28 (0.76*) per year & \cite{hollingsworth_hiv1_2008} \\
$D_P$ & Duration of primary infection & 1.23/12-6/12 (0.25*) years & \cite{hollingsworth_hiv1_2008} \\
$D_D$ & Duration of high transmission disease stage & 4.81/12-14/12 (0.75*) years & \cite{hollingsworth_hiv1_2008} \\
$\tsub{\beta}{max}$ & Maximum rate of transmission during asymptomatic stage & 0.317 per year & \cite{shirreff_transmission_2011} \\
$\beta_{50}$ & SPVL at which infectiousness is half maximum & 13938 copies per ml & \cite{shirreff_transmission_2011} \\
$\beta_k$ & Hill coefficient: steepness of increase in infectiousness as a function of SPVL & 1.02 & \cite{shirreff_transmission_2011} \\
$\tsub{D}{max}$ & Duration of primary infection & 25.4 years & \cite{shirreff_transmission_2011} \\
$D_{50}$ & SPVL at which duration of asymptomatic infection is half maximum & 3058 copies per ml & \cite{shirreff_transmission_2011} \\
$D_{k}$ & Hill coefficient: steepness of decrease in duration as a function of SPVL & 0.41 & \cite{shirreff_transmission_2011} \\
$\sigma_M$ & Mutation standard deviation of $\log_{10}$ SPVL & 0.12 & \cite{shirreff_transmission_2011} \\
$\tsub{\alpha}{min}$ & Minimum $\log_{10}$ SPVL & 2 & \cite{shirreff_transmission_2011}\\
$\tsub{\alpha}{max}$ & Maximum $\log_{10}$ SPVL & 7 & \cite{shirreff_transmission_2011}\\
$n$ & Number of strains & 21 (51*) & Assumption\\
$\mu$ & Mean number of non-cohabiting sexual partners & 0.103 - 1.206 & \cite{omori2015dynamics}\\
$\kappa$ & Squared coefficient of variation of number of non-cohabiting sexual partners & 0.01 - 100 & Assumption\\
\hline
\end{tabular}
\label{table:parmsTable}
\end{table}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig2.pdf}
\caption{{\bf Envelopes of virulence trajectories (population mean \Lspvl) under all models.}
All models were run until $t=4000~\textrm{years}$ using parameters from 1000 Latin hypercube samples to illustrate the range of dynamics each model can exhibit. Envelopes contain mid 95\% trajectories, while lines within the envelopes show mean trajectories. Truncated series (up to year 800) are shown here.}
\label{fig:virtraj}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig3.pdf}
\caption{{\bf Univariate distributions of summary statistics.}
For every simulation, three summary statistics were obtained: peak time, maximum mean \Lspvl, and equilibrium mean \Lspvl. 
Peak time is the time at which a virulence trajectory achieves its maximum value (maximum mean \Lspvl).
Equilibrium mean \Lspvl\ is taken by taking the mean \Lspvl\ value at $t = 4000$.
Minimum mean progression time is obtained by applying the hill function to maximum mean \Lspvl\ of each run.
The distribution of equilibrium mean \Lspvl\
(lower left panel) for the random mixing model is very narrow, and has been replaced by a point in order to preserve the vertical axis scaling.}
\label{fig:unidist}
\end{figure}

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig4.pdf}
\caption{{\bf Pairs plot: bivariate relationships among summary statistics for each model structure.}
Three summary statistics obtained from each simulation have been plotted against each other to visualize the relationship among the summary statistics and to allow for clearer comparison between models.
Surprisingly, the implicit model, an approximation for instantaneous partnership formation model (instswitch), shows an almost identical trend with a model that has pair formation dynamics (pairform).
Dashed line in equilibrium vs. peak virulence plot shows 1:1 line, where equilibrium and peak virulence are equal.
100 values were sampled from each model to allow for clearer distinction between the models}
\label{fig:pairplot}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{../figures/fig5.pdf}
\caption{{\bf Sensitivity plot.}
For most parameters in the Latin hypercube samples, the figure shows the distribution (points) and trend (smooth line) of the summary statistic as a function of the parameter value, some of which have adjusted to achieve the standard initial epidemic growth rate (see Methods).}
Opposing trends in summary statistics with respect to partnership dissolution rate, $c$, (third column) between the model with and without extra-pair contact illustrates that similar model with different assumptions may yield opposing predictions.
\label{fig:plot_sens}
\end{figure}
